Labeled object representation and manipulation with ITK

### 2007-04-27 article/Article.tex
more and more of the article
darcs-hash:20070427140655-2fc9d-c12abd5e7f427a2f5afdd2e11cdab5dd38fc8b3f.gz

Richard Beare has recently introduced a new filter to efficiently labelize the connected component with ITK, and has also proposed to use the run-length encoding used in that filter to implement some of the most useful binary mathematical morphology operators: the opening by attribute.
Following that idea, and after have searched a way to use the ITK's spatial objects for this task, a new set of classes have been developed to represent and manipulate the labeled images and the objects within them in ITK.
Those new classes have been used to implement several labeled images manipulation based on object attributes, as well as the binary specialization of some mathematical morphology filter already include in ITK, and not related to the attribute of the objects.

All the source codes are provided, as well as a full set of test and several usage example of the new classes.

@@

@@

@@

@@

@@

@@

>>>>

Richard Beare has recently introduced a new filter to efficiently labelize the
connected component with ITK, and has also proposed to use the run-length
encoding used in that filter to implement some of the most useful binary
mathematical morphology operators: the opening by attribute.
Following that idea, and after have searched a way to use the ITK's spatial
objects for this task, a new set of classes have been developed to represent and
manipulate the labeled images and the objects within them in ITK.
Those new classes have been used to implement several labeled images
manipulation based on object attributes, as well as the binary specialization of
some mathematical morphology filter already include in ITK, and not related to
the attribute of the objects. With those last filters, this contribution comes
with 49 new classes, and should greatly enhance the binary mathematical
morphology in ITK.

All the source codes are provided, as well as a full set of test and several
usage examples of the new classes.

@@

Once the manipulation of the objects is done, it can be useful to go back to a
more classic {\em itk::Image}. Several classes are provided to do that:
\begin{itemize}
  \item The {\em itk::LabelCollectionImageToLabelImageFilter} class simply
convert a {\em itk::LabelCollectionImage} to a labeled image stored in a {\em
itk::Image}.
  \item The {\em itk::LabelCollectionImageToBinaryImageFilter} put all the
objects in the foreground of a binary image stored in a {itk::Image}. It is
intended to be used with an image produced by the {\em
itk::BinaryImageToLabelCollectionImageFilter}. The background values of the
original image can also be restored by this filter.
  \item The {\em itk::LabelCollectionImageToMaskImageFilter} class can be used
to mask an image with the objects of the {\em itk::LabelCollectionImage}.
  \item Finally, {\em itk::LabelCollectionImageToAttributeImageFilter} produce
an {\em itk::Image} with the value of the attribute of the objects of the {\em
itk::LabelCollectionImage}. This filter is mostly useful to have a global view
of the attribute values in the image.
\end{itemize}

\section{Available attributes}

\subsection{Shape}

\begin{itemize}
  \item {\em Size}
  \item {\em PhysicalSize}
  \item {\em Centroid}
  \item {\em Region}
  \item {\em RegionElongation}
  \item {\em SizeRegionRatio}
  \item {\em SizeOnBorder}
  \item {\em FeretDiameter}
%   \item {\em Perimeter}
\end{itemize}

\subsection{Statistics}

\begin{itemize}
  \item {\em Minimum}
  \item {\em Maximum}
  \item {\em Mean}
  \item {\em Sum}
  \item {\em Sigma}
  \item {\em Variance}
  \item {\em Median}
  \item {\em MaximumIndex}
  \item {\em MinimumIndex}
  \item {\em CenterOfGravity}
%   \item {\em CentralMoments}
%   \item {\em PrincipalMoments}
%   \item {\em PrincipalAxes}
  \item {\em Kurtosis}
  \item {\em Skewness}
\end{itemize}

@@

The general view of the previous section show a very common way to use those
classes. To make easier to use, some prebuilt classes have been made, to perform
the mini-pipeline:
\begin{itemize}
  \item creation of the {\em itk::LabelCollectionImage} from an {\em
itk::Image},
  \item filtering of the {\em itk::LabelCollectionImage},
  \item creation of an {\em itk::Image} from the filtered {\em
itk::LabelCollectionImage},
\end{itemize}
with a specific attribute.

Those filters have been made for binary, and labeled images.

@@

\begin{itemize}
  \item {\em itk::BinaryAttributeKeepNObjectsImageFilter}
  \item {\em itk::BinaryAttributeOpeningImageFilter}
  \item {\em itk::BinaryShapeKeepNObjectsImageFilter}
  \item {\em itk::BinaryShapeOpeningImageFilter}
  \item {\em itk::BinaryStatisticsKeepNObjectsImageFilter}
  \item {\em itk::BinaryStatisticsOpeningImageFilter}
\end{itemize}

@@

\begin{itemize}
  \item {\em itk::LabelAttributeKeepNObjectsImageFilter}
  \item {\em itk::LabelAttributeOpeningImageFilter}
  \item {\em itk::LabelShapeKeepNObjectsImageFilter}
  \item {\em itk::LabelShapeOpeningImageFilter}
  \item {\em itk::LabelStatisticsKeepNObjectsImageFilter}
  \item {\em itk::LabelStatisticsOpeningImageFilter}
  \item {\em itk::ShapeRelabelImageFilter}
  \item {\em itk::StatisticsRelabelImageFilter}
\end{itemize}

@@

\begin{itemize}
  \item {\em itk::BinaryClosingByReconstructionImageFilter}
  \item {\em itk::BinaryFillholeImageFilter}
  \item {\em itk::BinaryGrindPeakImageFilter}
  \item {\em itk::BinaryOpeningByReconstructionImageFilter}
  \item {\em itk::BinaryReconstructionByDilationImageFilter}
  \item {\em itk::BinaryReconstructionByErosionImageFilter}
\end{itemize}

@@

\section{Wrappers support}

All the classes provided with that article, excepted the most generic ones made
to help the developer to implement some new features, can be used with WrapITK,
and have been fully tested with python.
